---
title: "Supplemental Analysis: GPT-2 Surprisal as Predictability Measure"
author: ""
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(brms)
library(mgcv)
library(mgcViz)
library(ggpubr)
library(tidybayes)
library(ggdist)
library(rlist)
library(plotly)
library(future)
```

## Overview

This document replicates the predictability analyses from the main paper, replacing the n-gram-based log-odds-ratio predictability measure with GPT-2 surprisal. Surprisal is defined as $-\log_e P(\textit{up} \mid \text{context})$, where context consists of all words preceding *up* in the sentence. Higher surprisal indicates that *up* is less predictable given the preceding context (i.e., surprisal is the inverse of predictability).

## GPT-2 Surprisal Computation

GPT-2 surprisal values are computed by `compute_gpt2_surprisal.py` (in this directory) and saved to `../Data/gpt2_surprisal.csv`. Run that script before knitting this document:

```
python compute_gpt2_surprisal.py
```

## Data Preparation

```{r load-data, message=FALSE}
gam_data = read_csv('../Data/gam_data.csv')
gpt2_surprisal = read_csv('../Data/gpt2_surprisal.csv')

gam_data = gam_data %>%
  left_join(gpt2_surprisal %>% select(Item, gpt2_surprisal), by = 'Item') %>%
  filter(!is.na(gpt2_surprisal))

gam_data$participant = as.factor(gam_data$participant)
gam_data$Item        = as.factor(gam_data$Item)
gam_data$Type        = as.factor(gam_data$Type)
gam_data$PhrasalVerb = as.factor(gam_data$PhrasalVerb)

cat("Items with GPT-2 surprisal:", length(unique(gam_data$Item)), "\n")
cat("Surprisal range: [", round(min(gam_data$gpt2_surprisal), 2), ",",
    round(max(gam_data$gpt2_surprisal), 2), "]\n")
```

## GAM Models

### Model 1: Interaction Between GPT-2 Surprisal and Frequency

This model replicates the model from Equation 1 of the paper, replacing `log_predic` with `gpt2_surprisal`.

$$
\begin{aligned}
\log(RT) & \sim te(\text{GPT-2 Surprisal}, \text{Frequency}) + \text{Duration} + s(\text{participant}, bs = \text{`}re\text{'}) \\
& + s(\text{Item}, bs = \text{`}re\text{'}) + s(\text{trial}, bs = \text{`}re\text{'}) \\
& + s(\text{GPT-2 Surprisal}, \text{Frequency}, \text{trial}, \text{participant}, bs = \text{`}re\text{'})
\end{aligned}
$$

```{r mod-gam1-gpt2, message=FALSE, eval=FALSE}
mod_gam1_gpt2 = bam(log_rt ~
                      te(gpt2_surprisal, log_freq) + Duration +
                      s(trial, bs = 're') +
                      s(participant, bs = 're') +
                      s(Item, bs = 're') +
                      s(gpt2_surprisal, log_freq, trial, participant, bs = 're'),
                    data = gam_data)

summary(mod_gam1_gpt2)
saveRDS(mod_gam1_gpt2, '../Models/mod_gam1_gpt2.rds')
```

```{r mod-gam1-gpt2-plot, message=FALSE, warning=FALSE, eval=FALSE}
surp_seq  = seq(min(gam_data$gpt2_surprisal), max(gam_data$gpt2_surprisal), length.out = 50)
freq_seq  = seq(min(gam_data$log_freq),        max(gam_data$log_freq),        length.out = 50)

pred_grid = expand.grid(gpt2_surprisal = surp_seq, log_freq = freq_seq) %>%
  mutate(Duration    = 0,
         trial       = gam_data$trial[1],
         participant = gam_data$participant[1],
         Item        = gam_data$Item[1])

preds = predict(mod_gam1_gpt2, newdata = pred_grid,
                exclude = c('s(trial)', 's(participant)', 's(Item)',
                            's(gpt2_surprisal,log_freq,trial,participant)'))

z_mat = matrix(preds, nrow = length(surp_seq), ncol = length(freq_seq))

plot_ly(x = surp_seq, y = freq_seq, z = z_mat,
        type = 'surface', colorscale = 'Viridis') %>%
  layout(scene = list(
    xaxis = list(title = 'GPT-2 Surprisal'),
    yaxis = list(title = 'Log Frequency'),
    zaxis = list(title = 'Predicted Log RT')
  ))
```

```{r mod-gam1-gpt2-slices, message=FALSE, warning=FALSE, eval=FALSE}
re_exclude_gam1 = c('s(trial)', 's(participant)', 's(Item)',
                    's(gpt2_surprisal,log_freq,trial,participant)')

slice_surp = tibble(
  gpt2_surprisal = seq(min(gam_data$gpt2_surprisal), max(gam_data$gpt2_surprisal), length.out = 200),
  log_freq       = mean(gam_data$log_freq),
  Duration       = 0,
  trial          = gam_data$trial[1],
  participant    = gam_data$participant[1],
  Item           = gam_data$Item[1]
)
preds_s = predict(mod_gam1_gpt2, newdata = slice_surp, se.fit = TRUE, exclude = re_exclude_gam1)
slice_surp = slice_surp %>% mutate(fit = preds_s$fit, se = preds_s$se.fit)

slice_freq = tibble(
  log_freq       = seq(min(gam_data$log_freq), max(gam_data$log_freq), length.out = 200),
  gpt2_surprisal = mean(gam_data$gpt2_surprisal),
  Duration       = 0,
  trial          = gam_data$trial[1],
  participant    = gam_data$participant[1],
  Item           = gam_data$Item[1]
)
preds_f = predict(mod_gam1_gpt2, newdata = slice_freq, se.fit = TRUE, exclude = re_exclude_gam1)
slice_freq = slice_freq %>% mutate(fit = preds_f$fit, se = preds_f$se.fit)

p1_surp = ggplot(slice_surp, aes(x = gpt2_surprisal, y = fit)) +
  geom_ribbon(aes(ymin = fit - se, ymax = fit + se), alpha = 0.3) +
  geom_line() +
  xlab('GPT-2 Surprisal\n(frequency at mean)') + ylab('Predicted Log RT') +
  theme_bw()

p1_freq = ggplot(slice_freq, aes(x = log_freq, y = fit)) +
  geom_ribbon(aes(ymin = fit - se, ymax = fit + se), alpha = 0.3) +
  geom_line() +
  xlab('Log Frequency\n(surprisal at mean)') + ylab('Predicted Log RT') +
  theme_bw()

ggarrange(p1_surp, p1_freq, nrow = 1)

rm(mod_gam1_gpt2, slice_surp, slice_freq, preds_s, preds_f, p1_surp, p1_freq); gc()
```

### Model 2: Phrasal vs. Non-Phrasal Verbs

This model replicates the model from Equation 2 of the paper, replacing `log_predic` with `gpt2_surprisal`.

$$
\begin{aligned}
\log(RT) & \sim te(\text{GPT-2 Surprisal}, \text{Frequency}, \text{by} = \text{PhrasalVerb}) + \text{Duration} \\
& + s(\text{participant}, bs = \text{`}re\text{'}) + s(\text{Item}, bs = \text{`}re\text{'}) + s(\text{trial}, bs = \text{`}re\text{'}) \\
& + s(\text{GPT-2 Surprisal}, \text{Frequency}, \text{trial}, \text{participant}, bs = \text{`}re\text{'})
\end{aligned}
$$

```{r mod-gam-phrasal-gpt2, message=FALSE, eval=FALSE}
mod_gam_phrasal_nonphrasal_gpt2 = bam(log_rt ~
                                        PhrasalVerb +
                                        te(gpt2_surprisal, log_freq, by = PhrasalVerb) + Duration +
                                        s(trial, bs = 're') +
                                        s(participant, bs = 're') +
                                        s(Item, bs = 're') +
                                        s(gpt2_surprisal, log_freq, trial, participant, bs = 're'),
                                      data = gam_data)

summary(mod_gam_phrasal_nonphrasal_gpt2)
saveRDS(mod_gam_phrasal_nonphrasal_gpt2, '../Models/mod_gam_phrasal_nonphrasal_gpt2.rds')
```

```{r mod-gam-phrasal-gpt2-plot, message=FALSE, warning=FALSE, eval=FALSE}
pred_grid_nonphrasal = pred_grid %>%
  mutate(PhrasalVerb = factor('nonphrasal', levels = levels(gam_data$PhrasalVerb)))

pred_grid_phrasal = pred_grid %>%
  mutate(PhrasalVerb = factor('phrasal', levels = levels(gam_data$PhrasalVerb)))

re_exclude = c('s(trial)', 's(participant)', 's(Item)',
               's(gpt2_surprisal,log_freq,trial,participant)')

preds_nonphrasal = predict(mod_gam_phrasal_nonphrasal_gpt2,
                           newdata = pred_grid_nonphrasal, exclude = re_exclude)
preds_phrasal    = predict(mod_gam_phrasal_nonphrasal_gpt2,
                           newdata = pred_grid_phrasal,    exclude = re_exclude)

z_nonphrasal = matrix(preds_nonphrasal, nrow = length(surp_seq), ncol = length(freq_seq))
z_phrasal    = matrix(preds_phrasal,    nrow = length(surp_seq), ncol = length(freq_seq))

scene_spec = list(xaxis = list(title = 'GPT-2 Surprisal'),
                  yaxis = list(title = 'Log Frequency'),
                  zaxis = list(title = 'Predicted Log RT'))

p_nonphrasal = plot_ly(x = surp_seq, y = freq_seq, z = z_nonphrasal,
                       type = 'surface', colorscale = 'Viridis', showscale = FALSE) %>%
  layout(title = 'Non-phrasal verbs', scene = scene_spec)

p_phrasal = plot_ly(x = surp_seq, y = freq_seq, z = z_phrasal,
                    type = 'surface', colorscale = 'Viridis', showscale = FALSE) %>%
  layout(title = 'Phrasal verbs', scene = scene_spec)

p_nonphrasal
p_phrasal
```

```{r mod-gam-phrasal-gpt2-slices, message=FALSE, warning=FALSE, eval=FALSE}
conditions = c('nonphrasal', 'phrasal')

slice_plots = lapply(conditions, function(cond) {
  base = tibble(Duration    = 0,
                trial       = gam_data$trial[1],
                participant = gam_data$participant[1],
                Item        = gam_data$Item[1],
                PhrasalVerb = factor(cond, levels = levels(gam_data$PhrasalVerb)))

  s_data = base %>%
    slice(rep(1, 200)) %>%
    mutate(gpt2_surprisal = seq(min(gam_data$gpt2_surprisal), max(gam_data$gpt2_surprisal), length.out = 200),
           log_freq       = mean(gam_data$log_freq))
  ps = predict(mod_gam_phrasal_nonphrasal_gpt2, newdata = s_data, se.fit = TRUE, exclude = re_exclude)
  s_data = s_data %>% mutate(fit = ps$fit, se = ps$se.fit)

  f_data = base %>%
    slice(rep(1, 200)) %>%
    mutate(log_freq       = seq(min(gam_data$log_freq), max(gam_data$log_freq), length.out = 200),
           gpt2_surprisal = mean(gam_data$gpt2_surprisal))
  pf = predict(mod_gam_phrasal_nonphrasal_gpt2, newdata = f_data, se.fit = TRUE, exclude = re_exclude)
  f_data = f_data %>% mutate(fit = pf$fit, se = pf$se.fit)

  ps_plot = ggplot(s_data, aes(x = gpt2_surprisal, y = fit)) +
    geom_ribbon(aes(ymin = fit - se, ymax = fit + se), alpha = 0.3) +
    geom_line() +
    xlab('GPT-2 Surprisal\n(frequency at mean)') + ylab('Predicted Log RT') +
    ggtitle(paste0(tools::toTitleCase(cond), ' verbs')) +
    theme_bw()

  pf_plot = ggplot(f_data, aes(x = log_freq, y = fit)) +
    geom_ribbon(aes(ymin = fit - se, ymax = fit + se), alpha = 0.3) +
    geom_line() +
    xlab('Log Frequency\n(surprisal at mean)') + ylab('Predicted Log RT') +
    ggtitle(paste0(tools::toTitleCase(cond), ' verbs')) +
    theme_bw()

  list(surp = ps_plot, freq = pf_plot)
})

ggarrange(slice_plots[[1]]$surp, slice_plots[[1]]$freq,
          slice_plots[[2]]$surp, slice_plots[[2]]$freq,
          nrow = 2, ncol = 2)

rm(mod_gam_phrasal_nonphrasal_gpt2, pred_grid, pred_grid_nonphrasal, pred_grid_phrasal,
   slice_plots, preds_nonphrasal, preds_phrasal, z_nonphrasal, z_phrasal); gc()
```

### Model 3: Main Effects and Interaction (ti() terms)

This model replicates the model from Equation 3 of the paper, replacing `log_predic` with `gpt2_surprisal`.

$$
\begin{aligned}
\log(RT) & \sim ti(\text{Frequency}) + ti(\text{GPT-2 Surprisal}) + ti(\text{GPT-2 Surprisal}, \text{Frequency}) \\
& + \text{Duration} + s(\text{participant}, bs = \text{`}re\text{'}) + s(\text{Item}, bs = \text{`}re\text{'}) \\
& + s(\text{GPT-2 Surprisal}, \text{Frequency}, \text{participant}, bs = \text{`}re\text{'})
\end{aligned}
$$

```{r mod-bam-inter-gpt2, message=FALSE, eval=FALSE}
mod_bam_inter_gpt2 = bam(log_rt ~
                           ti(log_freq) + ti(gpt2_surprisal) + ti(gpt2_surprisal, log_freq) +
                           Duration +
                           s(participant, bs = 're') +
                           s(Item, bs = 're') +
                           s(gpt2_surprisal, log_freq, participant, bs = 're'),
                         method = 'REML',
                         data = gam_data)

summary(mod_bam_inter_gpt2)
saveRDS(mod_bam_inter_gpt2, '../Models/mod_bam_inter_gpt2.rds')
```

```{r mod-bam-inter-gpt2-plot, message=FALSE, warning=FALSE, eval=FALSE}
nd_base = tibble(Duration = 0, log_freq = 0, gpt2_surprisal = 0, participant = 0, Item = 0)

nd_surp = nd_base %>%
  slice(rep(1, 1000)) %>%
  mutate(gpt2_surprisal = seq(min(gam_data$gpt2_surprisal), max(gam_data$gpt2_surprisal), length.out = 1000))
nd_surp = predict(mod_bam_inter_gpt2, newdata = nd_surp, se = TRUE, terms = 'ti(gpt2_surprisal)') %>%
  as_tibble() %>% cbind(nd_surp)

nd_freq = nd_base %>%
  slice(rep(1, 1000)) %>%
  mutate(log_freq = seq(min(gam_data$log_freq), max(gam_data$log_freq), length.out = 1000))
nd_freq = predict(mod_bam_inter_gpt2, newdata = nd_freq, se = TRUE, terms = 'ti(log_freq)') %>%
  as_tibble() %>% cbind(nd_freq)

p3_surp = ggplot(nd_surp, aes(x = gpt2_surprisal, y = fit)) +
  geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit), alpha = 0.3) +
  geom_line() +
  xlab('GPT-2 Surprisal of "up"') + ylab('Predicted Effect on Recognition Time') +
  theme_bw()

p3_freq = ggplot(nd_freq, aes(x = log_freq, y = fit)) +
  geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit), alpha = 0.3) +
  geom_line() +
  xlab('Log Frequency') + ylab('Predicted Effect on Recognition Time') +
  theme_bw()

ggarrange(p3_surp, p3_freq, nrow = 1)

rm(mod_bam_inter_gpt2, nd_base, nd_surp, nd_freq, p3_surp, p3_freq); gc()
```

## Bayesian Quadratic Models

All Bayesian models in this document are launched simultaneously in parallel using the `future` package (8 workers, 4 chains each). Models are cached to disk via the `file =` argument and loaded on subsequent knits without re-running.

```{r brms-launch, message=FALSE, include=FALSE}
options(contrasts = c("contr.sum", "contr.sum"))

priors = c(prior(student_t(3, 0, 2.5), class = 'sd'),
           prior(student_t(3, 0, 2.5), class = 'sigma'),
           prior(student_t(3, 0, 2.5), class = 'Intercept'),
           prior(normal(0, 0.1),       class = 'b'))

priors_lkj = c(prior(student_t(3, 0, 2.5), class = 'sd'),
               prior(student_t(3, 0, 2.5), class = 'sigma'),
               prior(student_t(3, 0, 2.5), class = 'Intercept'),
               prior(normal(0, 0.1),       class = 'b'),
               prior(lkj(2),               class = 'cor'))

gam_data_orig = read_csv('../Data/gam_data.csv') %>%
  mutate(participant = as.factor(participant),
         Item        = as.factor(Item),
         Type        = as.factor(Type),
         PhrasalVerb = as.factor(PhrasalVerb))

log_p_up       = log(594602106 / 1064909866902)
gam_data_local = gam_data_orig %>%
  mutate(log_cond_prob = log(plogis(log_predic)),
         log_pmi       = log_cond_prob - log_p_up)

plan(multisession, workers = 7)

# GPT-2 model futures omitted — those sections are eval=FALSE

# --- LKJ robustness models ---
future_lkj_full = future({
  library(brms)
  options(mc.cores = 4)
  brm(
    log_rt ~ log_freq + log_predic + Duration + I(log_predic^2) + I(log_freq^2) +
      (1 + log_freq + log_predic + Duration + I(log_predic^2) + I(log_freq^2) | participant) +
      (1 | Item),
    data = gam_data_orig, warmup = 15000, iter = 30000, cores = 4, chains = 4,
    seed = 964, init = 0, control = list(adapt_delta = 0.99),
    prior = priors_lkj, file = '../Models/brms_quadratic_lkj'
  )
})

future_lkj_freq = future({
  library(brms)
  options(mc.cores = 4)
  brm(
    log_rt ~ log_freq + Duration + I(log_freq^2) +
      (1 + log_freq + Duration + I(log_freq^2) | participant) +
      (1 | Item),
    data = gam_data_orig, warmup = 8000, iter = 16000, cores = 4, chains = 4,
    seed = 964, init = 0, control = list(adapt_delta = 0.99),
    prior = priors_lkj, file = '../Models/brms_freq_quadratic_lkj'
  )
})

future_lkj_predic = future({
  library(brms)
  options(mc.cores = 4)
  brm(
    log_rt ~ log_predic + Duration + I(log_predic^2) +
      (1 + log_predic + Duration + I(log_predic^2) | participant) +
      (1 | Item),
    data = gam_data_orig, warmup = 5000, iter = 10000, cores = 4, chains = 4,
    seed = 964, init = 0, control = list(adapt_delta = 0.99),
    prior = priors_lkj, file = '../Models/brms_predic_quadratic_lkj'
  )
})

# --- Log conditional probability models ---
future_cp_full = future({
  library(brms)
  options(mc.cores = 4)
  brm(
    log_rt ~ log_freq + log_cond_prob + Duration +
      I(log_cond_prob^2) + I(log_freq^2) +
      (1 + log_freq + log_cond_prob + Duration +
         I(log_cond_prob^2) + I(log_freq^2) || participant) +
      (1 || Item),
    data = gam_data_local, warmup = 10000, iter = 20000, cores = 4, chains = 4,
    seed = 964, init = 0, control = list(adapt_delta = 0.99),
    prior = priors, file = '../Models/brms_cond_prob_quadratic'
  )
})

future_cp_only = future({
  library(brms)
  options(mc.cores = 4)
  brm(
    log_rt ~ log_cond_prob + Duration + I(log_cond_prob^2) +
      (1 + log_cond_prob + Duration + I(log_cond_prob^2) || participant) +
      (1 || Item),
    data = gam_data_local, warmup = 10000, iter = 20000, cores = 4, chains = 4,
    seed = 964, init = 0, control = list(adapt_delta = 0.99),
    prior = priors, file = '../Models/brms_cond_prob_only_quadratic'
  )
})

future_cp_lkj_full = future({
  library(brms)
  options(mc.cores = 4)
  brm(
    log_rt ~ log_freq + log_cond_prob + Duration +
      I(log_cond_prob^2) + I(log_freq^2) +
      (1 + log_freq + log_cond_prob + Duration +
         I(log_cond_prob^2) + I(log_freq^2) | participant) +
      (1 | Item),
    data = gam_data_local, warmup = 15000, iter = 30000, cores = 4, chains = 4,
    seed = 964, init = 0, control = list(adapt_delta = 0.99),
    prior = priors_lkj, file = '../Models/brms_cond_prob_quadratic_lkj'
  )
})

future_cp_lkj_only = future({
  library(brms)
  options(mc.cores = 4)
  brm(
    log_rt ~ log_cond_prob + Duration + I(log_cond_prob^2) +
      (1 + log_cond_prob + Duration + I(log_cond_prob^2) | participant) +
      (1 | Item),
    data = gam_data_local, warmup = 10000, iter = 20000, cores = 4, chains = 4,
    seed = 964, init = 0, control = list(adapt_delta = 0.99),
    prior = priors_lkj, file = '../Models/brms_cond_prob_only_quadratic_lkj'
  )
})
```

### Full Model: Frequency and GPT-2 Surprisal

This model replicates Equation 4 of the paper, replacing `log_predic` with `gpt2_surprisal`.

$$
\begin{aligned}
\log(RT) & \sim \log(\text{Frequency}) + \text{GPT-2 Surprisal} + \text{Duration} \\
& + \log(\text{Frequency})^2 + \text{GPT-2 Surprisal}^2 \\
& + (1 + \log(\text{Frequency}) + \text{GPT-2 Surprisal} + \log(\text{Frequency})^2 \\
& \quad + \text{GPT-2 Surprisal}^2 + \text{Duration} \mid\mid \text{Participant}) + (1 \mid\mid \text{Item})
\end{aligned}
$$

```{r brms-full-gpt2, message=FALSE, eval=FALSE}
brms_gpt2_quadratic_no_interaction = value(future_gpt2_full)

fixef(brms_gpt2_quadratic_no_interaction)
```

```{r brms-full-gpt2-posterior, eval=FALSE}
percent_greater_zero = data.frame(fixef(brms_gpt2_quadratic_no_interaction, summary = FALSE)) %>%
  pivot_longer(cols = everything(), names_to = 'beta_coefficient', values_to = 'estimate') %>%
  group_by(beta_coefficient) %>%
  summarize(pct_gt_zero = (sum(estimate > 0) / n()) * 100)

percent_greater_zero
```

```{r brms-full-gpt2-plot, message=FALSE, warning=FALSE, out.width='90%', fig.align='center', eval=FALSE}
brms_gpt2_plot = conditional_effects(brms_gpt2_quadratic_no_interaction)

freq_plot_gpt2 = plot(brms_gpt2_plot, plot = FALSE)[[1]] +
  ylab('Log Reaction Time') + xlab('Log Frequency') + theme_bw()

surprisal_plot_gpt2 = plot(brms_gpt2_plot, plot = FALSE)[[2]] +
  ylab('Log Reaction Time') + xlab('GPT-2 Surprisal of "up"') + theme_bw()

ggarrange(freq_plot_gpt2, surprisal_plot_gpt2, nrow = 2, ncol = 1)
```

```{r brms-full-gpt2-posterior-plot, message=FALSE, warning=FALSE, out.width='80%', fig.align='center', eval=FALSE}
beta_coef_labels = list.reverse(c('GPT-2 Surprisal', 'log-frequency', 'Intercept',
                                   'GPT-2 Surprisal^2', 'log-frequency^2'))
brms_gpt2_quadratic_no_interaction %>%
  gather_draws(b_Intercept, b_log_freq, b_gpt2_surprisal,
               b_Igpt2_surprisalE2, b_Ilog_freqE2) %>%
  ggplot(aes(x = .value, y = .variable)) +
  stat_halfeye() +
  xlab('Posterior Distribution') +
  ylab('Beta Coefficient') +
  scale_y_discrete(labels = beta_coef_labels) +
  theme_bw()

rm(brms_gpt2_quadratic_no_interaction, brms_gpt2_plot,
   freq_plot_gpt2, surprisal_plot_gpt2, percent_greater_zero); gc()
```

### GPT-2 Surprisal-Only Model

This model replicates Equation 5 of the paper (the predictability-only quadratic), replacing `log_predic` with `gpt2_surprisal`.

$$
\begin{aligned}
\log(RT) & \sim \text{GPT-2 Surprisal} + \text{Duration} + \text{GPT-2 Surprisal}^2 \\
& + (1 + \text{GPT-2 Surprisal} + \text{GPT-2 Surprisal}^2 + \text{Duration} \mid\mid \text{Participant}) + (1 \mid\mid \text{Item})
\end{aligned}
$$

```{r brms-surprisal-only, message=FALSE, eval=FALSE}
brms_gpt2_surprisal_quadratic = value(future_gpt2_only)

fixef(brms_gpt2_surprisal_quadratic)
```

```{r brms-surprisal-only-posterior, eval=FALSE}
post_samples_surprisal = as.data.frame(fixef(brms_gpt2_surprisal_quadratic, summary = FALSE)) %>%
  select(gpt2_surprisal, Igpt2_surprisalE2)

cat("% samples gpt2_surprisal^2 > 0:",
    round(sum(post_samples_surprisal$Igpt2_surprisalE2 > 0) /
            nrow(post_samples_surprisal) * 100, 1), "\n")

cat("% samples gpt2_surprisal > 0:",
    round(sum(post_samples_surprisal$gpt2_surprisal > 0) /
            nrow(post_samples_surprisal) * 100, 1), "\n")
```

```{r brms-surprisal-only-plot, message=FALSE, warning=FALSE, out.width='60%', fig.align='center', eval=FALSE}
brms_surprisal_only_plot = conditional_effects(brms_gpt2_surprisal_quadratic)

plot(brms_surprisal_only_plot, plot = FALSE)[[1]] +
  ylab('Log Reaction Time') +
  xlab('GPT-2 Surprisal of "up"') +
  theme_bw()
```

```{r brms-surprisal-only-posterior-plot, message=FALSE, warning=FALSE, out.width='80%', fig.align='center', eval=FALSE}
beta_coef_labels_surp = list.reverse(c('GPT-2 Surprisal', 'Intercept', 'GPT-2 Surprisal^2'))
brms_gpt2_surprisal_quadratic %>%
  gather_draws(b_Intercept, b_gpt2_surprisal, b_Igpt2_surprisalE2) %>%
  ggplot(aes(x = .value, y = .variable)) +
  stat_halfeye() +
  xlab('Posterior Distribution') +
  ylab('Beta Coefficient') +
  scale_y_discrete(labels = beta_coef_labels_surp) +
  theme_bw()

rm(brms_gpt2_surprisal_quadratic, brms_surprisal_only_plot,
   post_samples_surprisal, gam_data); gc()
```

# Robustness Check: Correlated Random Effects with LKJ Priors

A reviewer noted that removing random-effect correlations (via `||`) for computational tractability is suboptimal and that LKJ priors on the correlation parameters would be preferable. The following models replicate the three Bayesian quadratic regression models from the paper with two changes: (1) random effects are allowed to correlate (using `|` instead of `||`), and (2) an LKJ(2) prior is placed on the correlation matrix. LKJ(2) is a weakly regularizing prior that places slightly more mass on near-identity matrices than the uniform LKJ(1), which helps with convergence while remaining diffuse.

All other model specifications (fixed effects, remaining priors, iterations) are identical to the original models.

```{r lkj-priors}
# priors_lkj and gam_data_orig are defined in the brms-launch chunk above
options(contrasts = c("contr.sum", "contr.sum"))
```

## Full Model: Frequency and Predictability with LKJ Priors

```{r brms-lkj-full, message=FALSE}
brms_quadratic_lkj = value(future_lkj_full)

fixef(brms_quadratic_lkj)
```

```{r brms-lkj-full-posterior}
percent_greater_zero_lkj = data.frame(fixef(brms_quadratic_lkj, summary = FALSE)) %>%
  pivot_longer(cols = everything(), names_to = 'beta_coefficient', values_to = 'estimate') %>%
  group_by(beta_coefficient) %>%
  summarize(pct_gt_zero = (sum(estimate > 0) / n()) * 100)

percent_greater_zero_lkj
```

```{r brms-lkj-full-posterior-plot, message=FALSE, warning=FALSE, out.width='80%', fig.align='center'}
beta_coef_labels_lkj = list.reverse(c('log-predictability', 'log-frequency', 'Intercept',
                                       'log-predictability^2', 'log-frequency^2'))
brms_quadratic_lkj %>%
  gather_draws(b_Intercept, b_log_freq, b_log_predic, b_Ilog_predicE2, b_Ilog_freqE2) %>%
  ggplot(aes(x = .value, y = .variable)) +
  stat_halfeye() +
  xlab('Posterior Distribution') +
  ylab('Beta Coefficient') +
  scale_y_discrete(labels = beta_coef_labels_lkj) +
  theme_bw()

rm(brms_quadratic_lkj, percent_greater_zero_lkj); gc()
```

## Frequency-Only Model with LKJ Priors

```{r brms-lkj-freq, message=FALSE}
brms_freq_quadratic_lkj = value(future_lkj_freq)

fixef(brms_freq_quadratic_lkj)

post_samples_freq_lkj = as.data.frame(fixef(brms_freq_quadratic_lkj, summary = FALSE)) %>%
  select(log_freq, Ilog_freqE2)

cat("% samples log_freq^2 > 0:",
    round(sum(post_samples_freq_lkj$Ilog_freqE2 > 0) / nrow(post_samples_freq_lkj) * 100, 1), "\n")
cat("% samples log_freq > 0:",
    round(sum(post_samples_freq_lkj$log_freq > 0) / nrow(post_samples_freq_lkj) * 100, 1), "\n")

rm(brms_freq_quadratic_lkj, post_samples_freq_lkj); gc()
```

## Predictability-Only Model with LKJ Priors

```{r brms-lkj-predic, message=FALSE}
brms_predic_quadratic_lkj = value(future_lkj_predic)

fixef(brms_predic_quadratic_lkj)

post_samples_predic_lkj = as.data.frame(fixef(brms_predic_quadratic_lkj, summary = FALSE)) %>%
  select(log_predic, Ilog_predicE2)

cat("% samples log_predic^2 > 0:",
    round(sum(post_samples_predic_lkj$Ilog_predicE2 > 0) / nrow(post_samples_predic_lkj) * 100, 1), "\n")
cat("% samples log_predic > 0:",
    round(sum(post_samples_predic_lkj$log_predic > 0) / nrow(post_samples_predic_lkj) * 100, 1), "\n")

rm(brms_predic_quadratic_lkj, post_samples_predic_lkj); gc()
```

---

# Local Context Analyses: Log Conditional Probability and PMI

These analyses replicate the predictability analyses from the main paper, replacing the log-odds-ratio predictability measure with log conditional probability $\log P(\textit{up} \mid V)$ and pointwise mutual information (PMI). Both measures quantify the strength of the local bigram association between the verb and *up* in the corpus, independent of sentential context.

## Relationship Between Measures

The `Predictability` column in the dataset is the bigram odds ratio $\text{Odds}(up \mid V) = P(up \mid V) / (1 - P(up \mid V))$. The log conditional probability is obtained via the logistic transform of the log-odds:

$$\log P(up \mid V) = \log\!\left(\frac{\text{Odds}}{1 + \text{Odds}}\right)$$

PMI is defined as:

$$\text{PMI}(V,\, up) = \log \frac{P(V,\, up)}{P(V)\,P(up)} = \log P(up \mid V) - \log P(up)$$

Because $\log P(up)$ is a corpus-level constant (frequency of *up* = 594,602,106; corpus size = 1,064,909,866,902), PMI and $\log P(up \mid V)$ are perfectly collinear and produce identical regression results (same slopes, different intercept). Models below use $\log P(up \mid V)$; PMI values are reported for descriptive completeness.

## Data Preparation

```{r local-data-prep, message=FALSE}
# gam_data_local (with log_cond_prob and log_pmi) is built in the brms-launch chunk above
cat("log P(up) =", round(log_p_up, 4), "\n")
cat("PMI offset from log P(up|V):", round(-log_p_up, 4), "\n")
cat("log_cond_prob range: [",
    round(min(gam_data_local$log_cond_prob, na.rm = TRUE), 3), ",",
    round(max(gam_data_local$log_cond_prob, na.rm = TRUE), 3), "]\n")
cat("log_pmi range: [",
    round(min(gam_data_local$log_pmi, na.rm = TRUE), 3), ",",
    round(max(gam_data_local$log_pmi, na.rm = TRUE), 3), "]\n")
cat("Pearson r(log_cond_prob, log_predic):",
    round(cor(gam_data_local$log_cond_prob, gam_data_local$log_predic,
              use = "complete.obs"), 4), "\n")
cat("Pearson r(log_cond_prob, log_freq):",
    round(cor(gam_data_local$log_cond_prob, gam_data_local$log_freq,
              use = "complete.obs"), 4), "\n")
```

## GAM Models

### Model 1: Interaction Between Log Conditional Probability and Frequency

This model replicates the model from Equation 1 of the paper, replacing `log_predic` with `log_cond_prob`.

$$
\begin{aligned}
\log(RT) & \sim te(\log P(up \mid V),\, \text{Frequency}) + \text{Duration} + s(\text{participant}, bs = \text{`}re\text{'}) \\
& + s(\text{Item}, bs = \text{`}re\text{'}) + s(\text{trial}, bs = \text{`}re\text{'}) \\
& + s(\log P(up \mid V),\, \text{Frequency},\, \text{trial},\, \text{participant}, bs = \text{`}re\text{'})
\end{aligned}
$$

```{r mod-gam1-cond-prob, message=FALSE}
mod_gam1_cond_prob = bam(log_rt ~
                           te(log_cond_prob, log_freq) + Duration +
                           s(trial, bs = 're') +
                           s(participant, bs = 're') +
                           s(Item, bs = 're') +
                           s(log_cond_prob, log_freq, trial, participant, bs = 're'),
                         data = gam_data_local)

summary(mod_gam1_cond_prob)
saveRDS(mod_gam1_cond_prob, '../Models/mod_gam1_cond_prob.rds')
```

```{r mod-gam1-cond-prob-plot, message=FALSE, warning=FALSE}
cond_seq  = seq(min(gam_data_local$log_cond_prob), max(gam_data_local$log_cond_prob), length.out = 50)
freq_seq2 = seq(min(gam_data_local$log_freq),       max(gam_data_local$log_freq),       length.out = 50)

pred_grid_cp = expand.grid(log_cond_prob = cond_seq, log_freq = freq_seq2) %>%
  mutate(Duration    = 0,
         trial       = gam_data_local$trial[1],
         participant = gam_data_local$participant[1],
         Item        = gam_data_local$Item[1])

preds_cp = predict(mod_gam1_cond_prob, newdata = pred_grid_cp,
                   exclude = c('s(trial)', 's(participant)', 's(Item)',
                               's(log_cond_prob,log_freq,trial,participant)'))

z_mat_cp = matrix(preds_cp, nrow = length(cond_seq), ncol = length(freq_seq2))

plot_ly(x = cond_seq, y = freq_seq2, z = z_mat_cp,
        type = 'surface', colorscale = 'Viridis') %>%
  layout(scene = list(
    xaxis = list(title = 'Log P(up|V)'),
    yaxis = list(title = 'Log Frequency'),
    zaxis = list(title = 'Predicted Log RT')
  ))
```

```{r mod-gam1-cond-prob-slices, message=FALSE, warning=FALSE}
re_excl_cp1 = c('s(trial)', 's(participant)', 's(Item)',
                's(log_cond_prob,log_freq,trial,participant)')

slice_cp = tibble(
  log_cond_prob = seq(min(gam_data_local$log_cond_prob), max(gam_data_local$log_cond_prob), length.out = 200),
  log_freq      = mean(gam_data_local$log_freq),
  Duration      = 0,
  trial         = gam_data_local$trial[1],
  participant   = gam_data_local$participant[1],
  Item          = gam_data_local$Item[1]
)
preds_cp_s = predict(mod_gam1_cond_prob, newdata = slice_cp, se.fit = TRUE, exclude = re_excl_cp1)
slice_cp = slice_cp %>% mutate(fit = preds_cp_s$fit, se = preds_cp_s$se.fit)

slice_freq_cp = tibble(
  log_freq      = seq(min(gam_data_local$log_freq), max(gam_data_local$log_freq), length.out = 200),
  log_cond_prob = mean(gam_data_local$log_cond_prob),
  Duration      = 0,
  trial         = gam_data_local$trial[1],
  participant   = gam_data_local$participant[1],
  Item          = gam_data_local$Item[1]
)
preds_cp_f = predict(mod_gam1_cond_prob, newdata = slice_freq_cp, se.fit = TRUE, exclude = re_excl_cp1)
slice_freq_cp = slice_freq_cp %>% mutate(fit = preds_cp_f$fit, se = preds_cp_f$se.fit)

p_cp_cond = ggplot(slice_cp, aes(x = log_cond_prob, y = fit)) +
  geom_ribbon(aes(ymin = fit - se, ymax = fit + se), alpha = 0.3) +
  geom_line() +
  xlab('Log P(up|V)\n(frequency at mean)') + ylab('Predicted Log RT') +
  theme_bw()

p_cp_freq = ggplot(slice_freq_cp, aes(x = log_freq, y = fit)) +
  geom_ribbon(aes(ymin = fit - se, ymax = fit + se), alpha = 0.3) +
  geom_line() +
  xlab('Log Frequency\n(log P(up|V) at mean)') + ylab('Predicted Log RT') +
  theme_bw()

ggarrange(p_cp_cond, p_cp_freq, nrow = 1)

rm(mod_gam1_cond_prob, slice_cp, slice_freq_cp, preds_cp_s, preds_cp_f, p_cp_cond, p_cp_freq); gc()
```

### Model 2: Phrasal vs. Non-Phrasal Verbs

This model replicates the model from Equation 2 of the paper, replacing `log_predic` with `log_cond_prob`.

```{r mod-gam-phrasal-cond-prob, message=FALSE}
mod_gam_phrasal_cond_prob = bam(log_rt ~
                                  PhrasalVerb +
                                  te(log_cond_prob, log_freq, by = PhrasalVerb) + Duration +
                                  s(trial, bs = 're') +
                                  s(participant, bs = 're') +
                                  s(Item, bs = 're') +
                                  s(log_cond_prob, log_freq, trial, participant, bs = 're'),
                                data = gam_data_local)

summary(mod_gam_phrasal_cond_prob)
saveRDS(mod_gam_phrasal_cond_prob, '../Models/mod_gam_phrasal_cond_prob.rds')
```

```{r mod-gam-phrasal-cond-prob-plot, message=FALSE, warning=FALSE}
pred_grid_cp_nonphrasal = pred_grid_cp %>%
  mutate(PhrasalVerb = factor('nonphrasal', levels = levels(gam_data_local$PhrasalVerb)))
pred_grid_cp_phrasal = pred_grid_cp %>%
  mutate(PhrasalVerb = factor('phrasal', levels = levels(gam_data_local$PhrasalVerb)))

re_excl_cp2 = c('s(trial)', 's(participant)', 's(Item)',
                's(log_cond_prob,log_freq,trial,participant)')

preds_cp_nonphrasal = predict(mod_gam_phrasal_cond_prob,
                              newdata = pred_grid_cp_nonphrasal, exclude = re_excl_cp2)
preds_cp_phrasal    = predict(mod_gam_phrasal_cond_prob,
                              newdata = pred_grid_cp_phrasal,    exclude = re_excl_cp2)

z_cp_nonphrasal = matrix(preds_cp_nonphrasal, nrow = length(cond_seq), ncol = length(freq_seq2))
z_cp_phrasal    = matrix(preds_cp_phrasal,    nrow = length(cond_seq), ncol = length(freq_seq2))

scene_cp = list(xaxis = list(title = 'Log P(up|V)'),
                yaxis = list(title = 'Log Frequency'),
                zaxis = list(title = 'Predicted Log RT'))

plot_ly(x = cond_seq, y = freq_seq2, z = z_cp_nonphrasal,
        type = 'surface', colorscale = 'Viridis', showscale = FALSE) %>%
  layout(title = 'Non-phrasal verbs', scene = scene_cp)

plot_ly(x = cond_seq, y = freq_seq2, z = z_cp_phrasal,
        type = 'surface', colorscale = 'Viridis', showscale = FALSE) %>%
  layout(title = 'Phrasal verbs', scene = scene_cp)
```

```{r mod-gam-phrasal-cond-prob-slices, message=FALSE, warning=FALSE}
conditions_cp = c('nonphrasal', 'phrasal')

slice_plots_cp = lapply(conditions_cp, function(cond) {
  base = tibble(Duration    = 0,
                trial       = gam_data_local$trial[1],
                participant = gam_data_local$participant[1],
                Item        = gam_data_local$Item[1],
                PhrasalVerb = factor(cond, levels = levels(gam_data_local$PhrasalVerb)))

  s_data = base %>%
    slice(rep(1, 200)) %>%
    mutate(log_cond_prob = seq(min(gam_data_local$log_cond_prob),
                               max(gam_data_local$log_cond_prob), length.out = 200),
           log_freq      = mean(gam_data_local$log_freq))
  ps = predict(mod_gam_phrasal_cond_prob, newdata = s_data, se.fit = TRUE, exclude = re_excl_cp2)
  s_data = s_data %>% mutate(fit = ps$fit, se = ps$se.fit)

  f_data = base %>%
    slice(rep(1, 200)) %>%
    mutate(log_freq      = seq(min(gam_data_local$log_freq),
                               max(gam_data_local$log_freq), length.out = 200),
           log_cond_prob = mean(gam_data_local$log_cond_prob))
  pf = predict(mod_gam_phrasal_cond_prob, newdata = f_data, se.fit = TRUE, exclude = re_excl_cp2)
  f_data = f_data %>% mutate(fit = pf$fit, se = pf$se.fit)

  ps_plot = ggplot(s_data, aes(x = log_cond_prob, y = fit)) +
    geom_ribbon(aes(ymin = fit - se, ymax = fit + se), alpha = 0.3) +
    geom_line() +
    xlab('Log P(up|V)\n(frequency at mean)') + ylab('Predicted Log RT') +
    ggtitle(paste0(tools::toTitleCase(cond), ' verbs')) +
    theme_bw()

  pf_plot = ggplot(f_data, aes(x = log_freq, y = fit)) +
    geom_ribbon(aes(ymin = fit - se, ymax = fit + se), alpha = 0.3) +
    geom_line() +
    xlab('Log Frequency\n(log P(up|V) at mean)') + ylab('Predicted Log RT') +
    ggtitle(paste0(tools::toTitleCase(cond), ' verbs')) +
    theme_bw()

  list(cond = ps_plot, freq = pf_plot)
})

ggarrange(slice_plots_cp[[1]]$cond, slice_plots_cp[[1]]$freq,
          slice_plots_cp[[2]]$cond, slice_plots_cp[[2]]$freq,
          nrow = 2, ncol = 2)

rm(mod_gam_phrasal_cond_prob, pred_grid_cp, pred_grid_cp_nonphrasal, pred_grid_cp_phrasal,
   slice_plots_cp, preds_cp_nonphrasal, preds_cp_phrasal, z_cp_nonphrasal, z_cp_phrasal); gc()
```

### Model 3: Main Effects and Interaction (ti() terms)

This model replicates the model from Equation 3 of the paper, replacing `log_predic` with `log_cond_prob`.

```{r mod-bam-inter-cond-prob, message=FALSE}
mod_bam_inter_cond_prob = bam(log_rt ~
                                ti(log_freq) + ti(log_cond_prob) + ti(log_cond_prob, log_freq) +
                                Duration +
                                s(participant, bs = 're') +
                                s(Item, bs = 're') +
                                s(log_cond_prob, log_freq, participant, bs = 're'),
                              method = 'REML',
                              data = gam_data_local)

summary(mod_bam_inter_cond_prob)
saveRDS(mod_bam_inter_cond_prob, '../Models/mod_bam_inter_cond_prob.rds')
```

```{r mod-bam-inter-cond-prob-plot, message=FALSE, warning=FALSE}
nd_base_cp = tibble(Duration = 0, log_freq = 0, log_cond_prob = 0, participant = 0, Item = 0)

nd_cond = nd_base_cp %>%
  slice(rep(1, 1000)) %>%
  mutate(log_cond_prob = seq(min(gam_data_local$log_cond_prob),
                             max(gam_data_local$log_cond_prob), length.out = 1000))
nd_cond = predict(mod_bam_inter_cond_prob, newdata = nd_cond,
                  se = TRUE, terms = 'ti(log_cond_prob)') %>%
  as_tibble() %>% cbind(nd_cond)

nd_freq_cp = nd_base_cp %>%
  slice(rep(1, 1000)) %>%
  mutate(log_freq = seq(min(gam_data_local$log_freq), max(gam_data_local$log_freq), length.out = 1000))
nd_freq_cp = predict(mod_bam_inter_cond_prob, newdata = nd_freq_cp,
                     se = TRUE, terms = 'ti(log_freq)') %>%
  as_tibble() %>% cbind(nd_freq_cp)

p3_cond = ggplot(nd_cond, aes(x = log_cond_prob, y = fit)) +
  geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit), alpha = 0.3) +
  geom_line() +
  xlab('Log P(up|V)') + ylab('Predicted Effect on Recognition Time') +
  theme_bw()

p3_freq_cp = ggplot(nd_freq_cp, aes(x = log_freq, y = fit)) +
  geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit), alpha = 0.3) +
  geom_line() +
  xlab('Log Frequency') + ylab('Predicted Effect on Recognition Time') +
  theme_bw()

ggarrange(p3_cond, p3_freq_cp, nrow = 1)

rm(mod_bam_inter_cond_prob, nd_base_cp, nd_cond, nd_freq_cp, p3_cond, p3_freq_cp); gc()
```

## Bayesian Quadratic Models

Priors are identical to those used in the main paper.

### Full Model: Frequency and Log Conditional Probability

This model replicates Equation 4 of the paper, replacing `log_predic` with `log_cond_prob`.

$$
\begin{aligned}
\log(RT) & \sim \log(\text{Frequency}) + \log P(up \mid V) + \text{Duration} \\
& + \log(\text{Frequency})^2 + \log P(up \mid V)^2 \\
& + (1 + \log(\text{Frequency}) + \log P(up \mid V) + \log(\text{Frequency})^2 \\
& \quad + \log P(up \mid V)^2 + \text{Duration} \mid\mid \text{Participant}) + (1 \mid\mid \text{Item})
\end{aligned}
$$

```{r brms-full-cond-prob, message=FALSE}
brms_cond_prob_quadratic = value(future_cp_full)

fixef(brms_cond_prob_quadratic)
```

```{r brms-full-cond-prob-posterior}
percent_greater_zero_cp = data.frame(fixef(brms_cond_prob_quadratic, summary = FALSE)) %>%
  pivot_longer(cols = everything(), names_to = 'beta_coefficient', values_to = 'estimate') %>%
  group_by(beta_coefficient) %>%
  summarize(pct_gt_zero = (sum(estimate > 0) / n()) * 100)

percent_greater_zero_cp
```

```{r brms-full-cond-prob-plot, message=FALSE, warning=FALSE, out.width='90%', fig.align='center'}
brms_cp_plot = conditional_effects(brms_cond_prob_quadratic)

freq_plot_cp = plot(brms_cp_plot, plot = FALSE)[[1]] +
  ylab('Log Reaction Time') + xlab('Log Frequency') + theme_bw()

cond_prob_plot = plot(brms_cp_plot, plot = FALSE)[[2]] +
  ylab('Log Reaction Time') + xlab('Log P(up|V)') + theme_bw()

ggarrange(freq_plot_cp, cond_prob_plot, nrow = 2, ncol = 1)
```

```{r brms-full-cond-prob-posterior-plot, message=FALSE, warning=FALSE, out.width='80%', fig.align='center'}
beta_coef_labels_cp = list.reverse(c('log P(up|V)', 'log-frequency', 'Intercept',
                                     'log P(up|V)^2', 'log-frequency^2'))
brms_cond_prob_quadratic %>%
  gather_draws(b_Intercept, b_log_freq, b_log_cond_prob,
               b_Ilog_cond_probE2, b_Ilog_freqE2) %>%
  ggplot(aes(x = .value, y = .variable)) +
  stat_halfeye() +
  xlab('Posterior Distribution') +
  ylab('Beta Coefficient') +
  scale_y_discrete(labels = beta_coef_labels_cp) +
  theme_bw()

rm(brms_cond_prob_quadratic, brms_cp_plot, percent_greater_zero_cp,
   freq_plot_cp, cond_prob_plot); gc()
```

### Log Conditional Probability-Only Model

This model replicates Equation 5 of the paper (the predictability-only quadratic), replacing `log_predic` with `log_cond_prob`.

$$
\begin{aligned}
\log(RT) & \sim \log P(up \mid V) + \text{Duration} + \log P(up \mid V)^2 \\
& + (1 + \log P(up \mid V) + \log P(up \mid V)^2 + \text{Duration} \mid\mid \text{Participant}) + (1 \mid\mid \text{Item})
\end{aligned}
$$

```{r brms-cond-prob-only, message=FALSE}
brms_cond_prob_only_quadratic = value(future_cp_only)

fixef(brms_cond_prob_only_quadratic)
```

```{r brms-cond-prob-only-posterior}
post_samples_cp_only = as.data.frame(fixef(brms_cond_prob_only_quadratic, summary = FALSE)) %>%
  select(log_cond_prob, Ilog_cond_probE2)

cat("% samples log_cond_prob^2 > 0:",
    round(sum(post_samples_cp_only$Ilog_cond_probE2 > 0) /
            nrow(post_samples_cp_only) * 100, 1), "\n")
cat("% samples log_cond_prob > 0:",
    round(sum(post_samples_cp_only$log_cond_prob > 0) /
            nrow(post_samples_cp_only) * 100, 1), "\n")
```

```{r brms-cond-prob-only-plot, message=FALSE, warning=FALSE, out.width='60%', fig.align='center'}
brms_cp_only_plot = conditional_effects(brms_cond_prob_only_quadratic)

plot(brms_cp_only_plot, plot = FALSE)[[1]] +
  ylab('Log Reaction Time') +
  xlab('Log P(up|V)') +
  theme_bw()
```

```{r brms-cond-prob-only-posterior-plot, message=FALSE, warning=FALSE, out.width='80%', fig.align='center'}
beta_coef_labels_cp_only = list.reverse(c('log P(up|V)', 'Intercept', 'log P(up|V)^2'))
brms_cond_prob_only_quadratic %>%
  gather_draws(b_Intercept, b_log_cond_prob, b_Ilog_cond_probE2) %>%
  ggplot(aes(x = .value, y = .variable)) +
  stat_halfeye() +
  xlab('Posterior Distribution') +
  ylab('Beta Coefficient') +
  scale_y_discrete(labels = beta_coef_labels_cp_only) +
  theme_bw()

rm(brms_cond_prob_only_quadratic, brms_cp_only_plot, post_samples_cp_only); gc()
```

## Robustness Check: Correlated Random Effects with LKJ Priors

These models replicate the two log conditional probability Bayesian models above with correlated random effects and an LKJ(2) prior on the correlation matrix, mirroring the robustness check for the original paper's models.

### Full Model: Frequency and Log Conditional Probability with LKJ Priors

```{r brms-cp-lkj-full, message=FALSE}
brms_cond_prob_quadratic_lkj = value(future_cp_lkj_full)

fixef(brms_cond_prob_quadratic_lkj)
```

```{r brms-cp-lkj-full-posterior}
percent_greater_zero_cp_lkj = data.frame(fixef(brms_cond_prob_quadratic_lkj, summary = FALSE)) %>%
  pivot_longer(cols = everything(), names_to = 'beta_coefficient', values_to = 'estimate') %>%
  group_by(beta_coefficient) %>%
  summarize(pct_gt_zero = (sum(estimate > 0) / n()) * 100)

percent_greater_zero_cp_lkj
```

```{r brms-cp-lkj-full-posterior-plot, message=FALSE, warning=FALSE, out.width='80%', fig.align='center'}
beta_coef_labels_cp_lkj = list.reverse(c('log P(up|V)', 'log-frequency', 'Intercept',
                                          'log P(up|V)^2', 'log-frequency^2'))
brms_cond_prob_quadratic_lkj %>%
  gather_draws(b_Intercept, b_log_freq, b_log_cond_prob,
               b_Ilog_cond_probE2, b_Ilog_freqE2) %>%
  ggplot(aes(x = .value, y = .variable)) +
  stat_halfeye() +
  xlab('Posterior Distribution') +
  ylab('Beta Coefficient') +
  scale_y_discrete(labels = beta_coef_labels_cp_lkj) +
  theme_bw()

rm(brms_cond_prob_quadratic_lkj, percent_greater_zero_cp_lkj); gc()
```

### Log Conditional Probability-Only Model with LKJ Priors

```{r brms-cp-lkj-only, message=FALSE}
brms_cond_prob_only_quadratic_lkj = value(future_cp_lkj_only)

fixef(brms_cond_prob_only_quadratic_lkj)
```

```{r brms-cp-lkj-only-posterior}
post_samples_cp_lkj_only = as.data.frame(fixef(brms_cond_prob_only_quadratic_lkj, summary = FALSE)) %>%
  select(log_cond_prob, Ilog_cond_probE2)

cat("% samples log_cond_prob^2 > 0:",
    round(sum(post_samples_cp_lkj_only$Ilog_cond_probE2 > 0) /
            nrow(post_samples_cp_lkj_only) * 100, 1), "\n")
cat("% samples log_cond_prob > 0:",
    round(sum(post_samples_cp_lkj_only$log_cond_prob > 0) /
            nrow(post_samples_cp_lkj_only) * 100, 1), "\n")
```

```{r brms-cp-lkj-only-posterior-plot, message=FALSE, warning=FALSE, out.width='80%', fig.align='center'}
beta_coef_labels_cp_lkj_only = list.reverse(c('log P(up|V)', 'Intercept', 'log P(up|V)^2'))
brms_cond_prob_only_quadratic_lkj %>%
  gather_draws(b_Intercept, b_log_cond_prob, b_Ilog_cond_probE2) %>%
  ggplot(aes(x = .value, y = .variable)) +
  stat_halfeye() +
  xlab('Posterior Distribution') +
  ylab('Beta Coefficient') +
  scale_y_discrete(labels = beta_coef_labels_cp_lkj_only) +
  theme_bw()

rm(brms_cond_prob_only_quadratic_lkj, post_samples_cp_lkj_only,
   gam_data_local, gam_data_orig); gc()
```